---
title: "COVID19 visualization (ECDC)"
author: "Stephan Michalik"
mail: "stephan.michalik@uni-greifswald.de"
output: 
  html_document:
    theme: yeti
    highlight: breezedark
    toc: true
    toc_depth: 2
    number_sections: true
    toc_float:
      collapsed: true
      smooth_scroll: true
      number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F,message=F,warning=F,fig.width=10,fig.height=6)
```

COVID 2019 data from European Centre for Disease Prevention and Control 



<!-- CALCULATIONS -->
```{r calculations}
# COVID 2019 data from European Centre for Disease Prevention and Control


#these libraries are necessary
library(R0)
library(shiny)
library(readxl)
library(httr)
library(tidyverse)
library(lubridate)
library(ggrepel)
library(plotly)
library(patchwork)
library(hrbrthemes)
library(DT)
library(helfRlein)
library(extrafont)
library(emojifont)
library(plotly)

#font 
# "Font Awesome 5 Free Solid" should be installed
# download and install from web: https://fontawesome.com/download

#colors setup
my_colors <- c("#5F4690","#1D6996","#38A6A5","#0F8554","#73AF48","#EDAD08","#E17C05","#CC503E","#94346E","#6F4070","#994E95","#666666")

#create the URL where the dataset is stored with automatic updates every day


#GET("https://www.ecdc.europa.eu/sites/default/files/documents/COVID-19-geographic-disbtribution-worldwide.xlsx", authenticate(":", ":", type="ntlm"), write_disk(tf <- tempfile(fileext = ".xlsx")))

GET("https://opendata.ecdc.europa.eu/covid19/casedistribution/csv", authenticate(":", ":", type="ntlm"), write_disk(tf <- tempfile(fileext = ".csv")))
#download the dataset from the website to a local temporary file

#read the Dataset sheet into “R”. The dataset will be called "data".
#data_in <- read_excel(tf)
data_in <- read_csv(tf)
# date format----

data_in <- data_in %>% mutate(dateRep = as.Date(dateRep, "%d/%m/%Y"))

#rename contries Territories----
data_in <- data_in %>% mutate(countriesAndTerritories= char_replace(countriesAndTerritories))

#write_file----
write_delim(x = data_in,path = "COVID_cases_deaths_reported_by_ECDC.tab",delim = "\t",col_names = T)

data_aggregated <- data_in %>% 
  group_by(countriesAndTerritories) %>% 
  summarise(sum_of_cases = sum(cases,na.rm = T),sum_of_deaths = sum(deaths,na.rm = T)) %>% 
  arrange(desc(sum_of_cases))




#sum up day for day cases and deaths----
data_in <- data_in %>%  mutate(day_start = dense_rank(dateRep))

data_in$day_sum_cases <- NA
data_in$day_sum_deaths <- NA

for(i in unique(data_in$geoId)){
  for(k in unique(data_in$day_start)){
    data_in$day_sum_cases[which(data_in$geoId==i & data_in$day_start==k)] <- 
      sum(data_in$cases[which(data_in$geoId==i & data_in$day_start<=k)])
    data_in$day_sum_deaths[which(data_in$geoId==i & data_in$day_start==k)] <- 
      sum(data_in$deaths[which(data_in$geoId==i & data_in$day_start<=k)])
  }
}

#generate current data
current_data_in <- data_in %>% 
                filter(dateRep==max(dateRep)) %>% 
                select(countriesAndTerritories,cases,deaths,day_sum_cases,day_sum_deaths) %>% 
                arrange(desc(day_sum_cases)) %>% 
                rename(total_cases = day_sum_cases,total_deaths = day_sum_deaths)


# global area plot----
global_cases_deaths_daily <- data_in %>% group_by(dateRep) %>% summarise(cases = sum(cases,na.rm = T), deaths = sum(deaths,na.rm = T)) %>%
gather(cases,deaths,key = "cases_or_deaths",value = "value")

global_cases_deaths_daily_plot<-ggplot(global_cases_deaths_daily,mapping = aes(dateRep,value,fill=cases_or_deaths))+
  geom_area()+
  scale_fill_brewer(palette = "Set1",direction = -1)+
  labs(title = "daily cases and deaths (whole world)",
       subtitle = "data from European Centre for Disease Prevention and Control",
       x = "date",
       fill = "",
       y = "daily cases/deaths")+
    theme_ipsum_rc(base_size = 18)+
  geom_label_repel(data = global_cases_deaths_daily %>% filter(dateRep==max(dateRep)),mapping = aes(label = format(value, nsmall=0, big.mark=",")),segment.colour = "black")


global_cases_deaths_overall<- data_in %>% group_by(dateRep) %>% summarise(cases = sum(day_sum_cases,na.rm = T), deaths = sum(day_sum_deaths,na.rm = T)) %>%
gather(cases,deaths,key = "cases_or_deaths",value = "value")

global_cases_deaths_overall_plot <- ggplot(data = global_cases_deaths_overall,mapping = aes(dateRep,value,fill=cases_or_deaths))+
  geom_area()+
  scale_fill_brewer(palette = "Set1",direction = -1)+
  labs(title = "cases and deaths (whole world) since 2019-12-31",
       subtitle = "data from European Centre for Disease Prevention and Control",
       x = "date",
       fill = "",
       y = "cases/deaths since 2019-12-31")+
    theme_ipsum_rc(base_size = 18)+
  geom_label_repel(data = global_cases_deaths_overall %>% filter(dateRep==max(dateRep)),mapping = aes(label = format(value, nsmall=0, big.mark=",")),segment.colour = "black")


# create valueBoxes----
##prepare a data object to render a summary graphic

current_global_overall <- global_cases_deaths_overall %>% filter(dateRep==max(dateRep))
current_global_daily <- global_cases_deaths_daily %>% filter(dateRep==max(dateRep))

figures <- 4
df <- data.frame(
    x = cumsum(c(2, rep(6.5, figures-1))),
    y = rep(2, figures),
    h = rep(4, figures),
    w = rep(6, figures))
df$key <- c("daily cases", "daily deaths","summed cases","summed deaths")
df$icon <- fontawesome(c("fa-hospital-o","fa-ban","fa-hospital-o","fa-ban"))
df$colour <- c("#006064","#BF360C","#01579B","#B71C1C")
df$text_colour <- c("#B2EBF2","#FFCCBC","#B3E5FC","#FFCDD2")
df$info <- c(current_global_daily$value[1],
             current_global_daily$value[2],
             current_global_overall$value[1],
             current_global_overall$value[2])
#valueBox display----
SummaryValueBoxes <- ggplot(df, aes(x, y, height = h, width = w, label = key)) +
    geom_tile(aes(fill=key)) +
  scale_fill_manual(values = df$colour)+
    geom_text(color = df$text_colour, hjust="left", nudge_y=-1.5, nudge_x=-2.6, size=5) +
    geom_text(label=format(df$info, nsmall=0, big.mark=","), size=10, color = "white", fontface = "bold", nudge_x=-2.6, hjust="left") +
  geom_text(label=df$icon, family="Font Awesome 5 Free Solid", colour="white", size=10, hjust="right", nudge_x=2.85, nudge_y=0.85) +
    coord_fixed() + 
    theme_void() +
    guides(fill = F)




#select  Top10 countries----
top10<- rbind(data_aggregated[1:10,],
              data_aggregated %>% filter(countriesAndTerritories=="Qatar"),
              data_aggregated %>% filter(countriesAndTerritories=="Sweden"))
data_filtered <- data_in %>% filter(countriesAndTerritories%in%top10$countriesAndTerritories)

#calculate date when 5th death was reported----
data_filtered <- data_filtered %>% group_by(geoId) %>% 
  mutate(days_since_5th_death = day_start-day_start[which(abs(day_sum_deaths-5)==min(abs(day_sum_deaths-5)))][1]) %>% 
  ungroup

#calculate date when 1000th case was reported----
data_filtered <- data_filtered %>% group_by(geoId) %>% 
  mutate(days_since_1000th_case = day_start-day_start[which(abs(day_sum_cases-1000)==min(abs(day_sum_cases-1000)))][1]) %>% 
  ungroup


#case plot----
cases_plot <- ggplot(data_filtered, aes(day_start, day_sum_cases, color=geoId)) +
  geom_point(aes(size = day_sum_deaths))+
  labs(title = "TOP10 cases - countries: cases of COVID19 ",
       subtitle = "data from European Centre for Disease Prevention and Control",
       size = "deaths since\n 2019-12-31",
       x = "days since 2019-12-31",
       y = "cumulated cases")+
  scale_color_manual(values = my_colors)+
  geom_label_repel(data_filtered %>% filter(day_start==max(data_filtered$day_start)),
                   mapping = aes(label=countriesAndTerritories))+
  theme_ipsum_rc(base_size = 18)+
  guides(color=F)

#deaths plot----
deaths_plot <- ggplot(data_filtered, aes(day_start, day_sum_deaths, color=geoId)) +
  geom_point(aes(size = day_sum_cases))+
  labs(title = "TOP10 cases - countries: deaths of COVID19 ",
       subtitle = "data from European Centre for Disease Prevention and Control",
       size = "cases since\n 2019-12-31",
       x = "days since 2019-12-31",
       y = "cumulated deaths")+
  scale_color_manual(values = my_colors)+
  geom_label_repel(data_filtered %>% filter(day_start==max(data_filtered$day_start)),
                   mapping = aes(label=countriesAndTerritories))+
  theme_ipsum_rc(base_size = 18)+
  guides(color=F)



#deaths plot 1000th reported----
deaths_plot_1000th_case <- ggplot(data_filtered %>% filter(days_since_1000th_case>=0), aes(days_since_1000th_case, day_sum_deaths, color=geoId)) +
  geom_point(aes(size = day_sum_cases))+
  labs(title = "TOP10 cases - countries: deaths of COVID19 ",
       subtitle = "data from European Centre for Disease Prevention and Control",
       size = "cases since\n 2019-12-31",
       x = "days since 1000th cases was reported in country",
       y = "cumulated deaths")+
  scale_color_manual(values = my_colors)+
  geom_label_repel(data_filtered %>% filter(day_start==max(data_filtered$day_start)),
                   mapping = aes(label=countriesAndTerritories),box.padding = 0.5)+
  theme_ipsum_rc(base_size = 18)+
  guides(color=F)


deaths_plot_1000th_case_facet <- ggplot(data_filtered %>% filter(days_since_1000th_case>=0), aes(days_since_1000th_case, day_sum_deaths, color=geoId)) +
  stat_smooth(method = "gam",fullrange = T,formula = y ~ s(x, bs = "cs"))+
  geom_point(aes(size = day_sum_cases,))+
  labs(title = "TOP10 cases - countries: deaths of COVID19 ",
       subtitle = "data from European Centre for Disease Prevention and Control",
       size = "cases since\n 2019-12-31",
       x = "days since 1000th cases was reported in country",
       y = "cumulated deaths", 
       caption ="Generalized additive models with integrated smoothness estimation were added -- y ~ s(x, bs = 'cs')")+
  scale_color_manual(values = my_colors)+
  theme_ipsum_rc(base_size = 18)+
  guides(color=F)+
  facet_wrap(~countriesAndTerritories)




#deaths per day aligned----

deaths_plot_1000th_case_per_day <- ggplot(data_filtered %>% filter(days_since_1000th_case>=0), aes(days_since_1000th_case, deaths, fill=geoId)) +
  geom_bar(stat="identity")+
  labs(title = "TOP10 cases - countries: deaths of COVID19 per day",
       subtitle = "data from European Centre for Disease Prevention and Control",
       size = "cases since\n 2019-12-31",
       x = "days since 1000th cases was reported in country",
       y = "deaths per day")+
  scale_fill_manual(values = my_colors)+
  theme_ipsum_rc(base_size = 18)+
  guides(fill=F)+
  facet_wrap(~countriesAndTerritories)



#cases per day aligned----
cases_plot_1000th_case_per_day <- ggplot(data_filtered %>% filter(days_since_1000th_case>=0), aes(days_since_1000th_case, cases, fill=geoId)) +
  geom_bar(stat="identity")+
  labs(title = "TOP10 cases - countries: cases of COVID19 per day",
       subtitle = "data from European Centre for Disease Prevention and Control",
       size = "cases since\n 2019-12-31",
       x = "days since 1000th cases was reported in country",
       y = "cases per day")+
  scale_fill_manual(values = my_colors)+
  theme_ipsum_rc(base_size = 18)+
  guides(fill=F)+
  facet_wrap(~countriesAndTerritories)

# add curves doubling times ----
#add curves
doubling_time <- seq(1,6)
time<- seq(0,60,by = 0.1)
initial_cases <- 1000

estimated_cases_doubling_time <- c()
for(i in 1:length(doubling_time)){
  
  estimated_cases<- initial_cases*exp(1)^(time*(log(2)/doubling_time[i])) 
  estimated_cases_doubling_time <- bind_rows(estimated_cases_doubling_time,
                                             tibble(days = time, 
                                                    cases = estimated_cases,
                                                    doubling_time = doubling_time[i]))
  
}




#plot plot 1000th case reported----
case_plot_1000th_case <- ggplot(data_filtered %>% filter(days_since_1000th_case>=0), aes(days_since_1000th_case, day_sum_cases, color=geoId)) +
  geom_line(data = estimated_cases_doubling_time,
            mapping = aes(days,cases,group = doubling_time,linetype=as.factor(doubling_time)),color="grey")+
  ylim(c(1000,max(data_filtered$day_sum_cases)*0.4+max(data_filtered$day_sum_cases)))+
    geom_label_repel(data= estimated_cases_doubling_time %>% 
                       filter(cases < max(data_filtered$day_sum_cases)*0.4+
                                max(data_filtered$day_sum_cases)) %>% 
                       group_by(doubling_time) %>% 
                       top_n(n = 1,wt = cases),
                     mapping = aes(x = days,
                                   y = cases,
                                   label=paste(doubling_time,"d",sep="")),
                     inherit.aes = F,
                     color="grey")+
  geom_point(aes(size = day_sum_deaths))+
  labs(title = "TOP10 cases - countries: cases of COVID19 ",
       subtitle = "data from European Centre for Disease Prevention and Control",
       size = "deaths since\n 2019-12-31",
       x = "days since 1000th case was reported in country",
       y = "cumulated cases")+
  scale_color_manual(values = my_colors)+
  geom_label_repel(data_filtered %>% filter(day_start==max(data_filtered$day_start)),
                   mapping = aes(label=countriesAndTerritories),box.padding = 0.5)+
  theme_ipsum_rc(base_size = 18)+
  guides(color=F)+
  guides(linetype=F)


case_plot_1000th_case_facet <- ggplot(data_filtered %>% filter(days_since_1000th_case>=0), aes(days_since_1000th_case, day_sum_cases, color=geoId)) +
  stat_smooth(method = "gam",fullrange = T,formula = y ~ s(x, bs = "cs"))+
  geom_line(data = estimated_cases_doubling_time,
            mapping = aes(days,cases,group = doubling_time,linetype=as.factor(doubling_time)),color="grey")+
  ylim(c(1000,max(data_filtered$day_sum_cases)*0.4+
           max(data_filtered$day_sum_cases)))+
    geom_label_repel(data= estimated_cases_doubling_time %>% 
                       filter(cases < max(data_filtered$day_sum_cases)*0.4+
                                max(data_filtered$day_sum_cases)) %>% 
                       group_by(doubling_time) %>% 
                       top_n(n = 1,wt = cases),
                     mapping = aes(x = days,
                                   y = cases,
                                   label=paste(doubling_time,"d",sep="")),
                     inherit.aes = F,
                     color="grey", size=3)+
  geom_point(aes(size = day_sum_deaths))+
  labs(title = "TOP10 cases - countries: cases of COVID19 ",
       subtitle = "data from European Centre for Disease Prevention and Control",
       size = "deaths since\n 2019-12-31",
       x = "days since 1000th case was reported in country",
       caption = "Generalized additive models with integrated smoothness estimation were added -- y ~ s(x, bs = 'cs')",
       y = "cumulated cases")+
  scale_color_manual(values = my_colors)+
  theme_ipsum_rc(base_size = 18)+
  guides(color=F)+
  guides(linetype=F)+
  facet_wrap(~countriesAndTerritories)




#calculate R0 Germany ----

Germany_cases_100 <- data_filtered %>% filter(countriesAndTerritories=="Germany" & days_since_1000th_case>=-8)
Germany_cases_100_vector <- Germany_cases_100$cases
names(Germany_cases_100_vector) <- Germany_cases_100$dateRep
mGT <- generation.time(type = "gamma",val =  c(4,2))
TD <- est.R0.TD(rev(Germany_cases_100_vector), mGT, nsim=10000,begin = 1,end = length(Germany_cases_100_vector)-1)## Results will include "most likely R(t)" (ie. the R(t)
R0_Germany <- tibble(dateRep = date(names(TD$R)),R0 = TD$R,lower_CI = TD$conf.int$lower,upper_CI = TD$conf.int$upper)
R0_Germany <- left_join(Germany_cases_100,R0_Germany,by="dateRep")
#plot
R0_Germany_plot<- ggplot(R0_Germany,aes(as_date(dateRep),R0))+
  geom_hline(yintercept = 1,linetype="dashed",color = "darkgrey")+
  geom_line(aes(as_date(dateRep),upper_CI), color="grey",linetype = "dotted")+
  geom_line(aes(as_date(dateRep),lower_CI), color="grey",linetype = "dotted")+
  geom_hline(yintercept = 0)+
  geom_point(color="grey",size=3)+
  geom_line(color="grey",size=1.5)+
  scale_y_continuous(breaks=seq(0,8,by = 0.5))+
  theme_ipsum_rc(base_size = 18)+
  labs(title = "Germany: R0",subtitle = "generation time (gamma distribution with a mean of 4 (SD=2)",x = "date")+
  geom_text_repel(data= R0_Germany %>% top_n(n = 20,wt = as_date(dateRep)),mapping = aes(label = paste(dateRep,"\nR0 = ",round(R0,2),sep="")),size=2)#

Germany_plot_daily <- ggplot(Germany_cases_100,aes(dateRep,cases))+
  geom_hline(yintercept = 0)+
   geom_point(color="grey",size=3)+
  geom_line(color="grey",size=1.5)+
  theme_ipsum_rc(base_size = 18)+
  labs(title = "Germany: daily cases",subtitle = "",x = "date")+
  geom_text_repel(Germany_cases_100 %>% top_n(n = 20,wt = as_date(dateRep)),mapping = aes(label = paste(dateRep,"\ncases = ",cases,sep="")),size=2)


#calculate R0 USA ----

USA_cases_100 <- data_filtered %>% filter(countriesAndTerritories=="United_States_of_America" & days_since_5th_death>=0)
USA_cases_100_vector <- USA_cases_100$cases
names(USA_cases_100_vector) <- USA_cases_100$dateRep
mGT <- generation.time(type = "gamma",val =  c(4,2))
TD_USA <- est.R0.TD(rev(USA_cases_100_vector), mGT, nsim=10000,begin = 1,end = length(USA_cases_100_vector)-1)## Results will include "most likely R(t)" (ie. the R(t)
R0_USA <- tibble(dateRep = date(names(TD_USA$R)),R0 = TD_USA$R,lower_CI = TD_USA$conf.int$lower,upper_CI = TD_USA$conf.int$upper)
R0_USA <- left_join(USA_cases_100,R0_USA,by="dateRep")
#plot
R0_USA_plot<- ggplot(R0_USA,aes(as_date(dateRep),R0))+
  geom_hline(yintercept = 1,linetype="dashed",color = "darkgrey")+
  geom_line(aes(as_date(dateRep),upper_CI), color="grey",linetype = "dotted")+
  geom_line(aes(as_date(dateRep),lower_CI), color="grey",linetype = "dotted")+
  geom_hline(yintercept = 0)+
  geom_point(color="grey",size=3)+
  geom_line(color="grey",size=1.5)+
  scale_y_continuous(breaks=seq(0,8,by = 0.5))+
  theme_ipsum_rc(base_size = 18)+
  labs(title = "USA: R0",subtitle = "generation time (gamma distribution with a mean of 4 (SD=2)",x = "date")+
  geom_text_repel(data = R0_USA%>% top_n(n = 20,wt = as_date(dateRep)),mapping = aes(label = paste(dateRep,"\nR0 = ",round(R0,2),sep="")),size=2)

USA_plot_daily <- ggplot(USA_cases_100,aes(dateRep,cases))+
  geom_hline(yintercept = 0)+
   geom_point(color="grey",size=3)+
  geom_line(color="grey",size=1.5)+
  theme_ipsum_rc(base_size = 18)+
  labs(title = "USA: daily cases",subtitle = "",x = "date")+
  geom_text_repel(mapping = aes(label = paste(dateRep,"\ncases = ",cases,sep="")),size=2)



#calculate R0 Sweden ----

Sweden_cases_100 <- data_filtered %>% filter(countriesAndTerritories=="Sweden" & days_since_1000th_case>=-9)
Sweden_cases_100_vector <- Sweden_cases_100$cases
names(Sweden_cases_100_vector) <- Sweden_cases_100$dateRep
mGT <- generation.time(type = "gamma",val =  c(4,2))
TD <- est.R0.TD(rev(Sweden_cases_100_vector), mGT, nsim=10000,begin = 1,end = length(Sweden_cases_100_vector)-1)## Results will include "most likely R(t)" (ie. the R(t)
R0_Sweden <- tibble(dateRep = date(names(TD$R)),R0 = TD$R,lower_CI = TD$conf.int$lower,upper_CI = TD$conf.int$upper)
R0_Sweden <- left_join(Sweden_cases_100,R0_Sweden,by="dateRep")
#plot
R0_Sweden_plot<- ggplot(R0_Sweden,aes(as_date(dateRep),R0))+
  geom_hline(yintercept = 1,linetype="dashed",color = "darkgrey")+
  geom_line(aes(as_date(dateRep),upper_CI), color="grey",linetype = "dotted")+
  geom_line(aes(as_date(dateRep),lower_CI), color="grey",linetype = "dotted")+
  geom_hline(yintercept = 0)+
  geom_point(color="grey",size=3)+
  geom_line(color="grey",size=1.5)+
  scale_y_continuous(breaks=seq(0,8,by = 0.5))+
  theme_ipsum_rc(base_size = 18)+
  labs(title = "Sweden: R0",subtitle = "generation time (gamma distribution with a mean of 4 (SD=2)",x = "date")+
  geom_text_repel(mapping = aes(label = paste(dateRep,"\nR0 = ",round(R0,2),sep="")),size=2)#

Sweden_plot_daily <- ggplot(Sweden_cases_100,aes(dateRep,cases))+
  geom_hline(yintercept = 0)+
   geom_point(color="grey",size=3)+
  geom_line(color="grey",size=1.5)+
  theme_ipsum_rc(base_size = 18)+
  labs(title = "Sweden: daily cases",subtitle = "",x = "date")+
  geom_text_repel(mapping = aes(label = paste(dateRep,"\ncases = ",cases,sep="")),size=2)



#calculate R0 Qatar ----

Qatar_cases_100 <- data_filtered %>% filter(countriesAndTerritories=="Qatar" & days_since_1000th_case>=-20)
Qatar_cases_100_vector <- Qatar_cases_100$cases
names(Qatar_cases_100_vector) <- Qatar_cases_100$dateRep
mGT <- generation.time(type = "gamma",val =  c(4,2))
TD <- est.R0.TD(rev(Qatar_cases_100_vector), mGT, nsim=10000,begin = 1,end = length(Qatar_cases_100_vector)-1)## Results will include "most likely R(t)" (ie. the R(t)
R0_Qatar <- tibble(dateRep = date(names(TD$R)),R0 = TD$R,lower_CI = TD$conf.int$lower,upper_CI = TD$conf.int$upper)
R0_Qatar <- left_join(Qatar_cases_100,R0_Qatar,by="dateRep")
#plot
R0_Qatar_plot<- ggplot(R0_Qatar,aes(as_date(dateRep),R0))+
  geom_hline(yintercept = 1,linetype="dashed",color = "darkgrey")+
  geom_line(aes(as_date(dateRep),upper_CI), color="grey",linetype = "dotted")+
  geom_line(aes(as_date(dateRep),lower_CI), color="grey",linetype = "dotted")+
  geom_hline(yintercept = 0)+
  geom_point(color="grey",size=3)+
  geom_line(color="grey",size=1.5)+
  scale_y_continuous(breaks=seq(0,8,by = 0.5))+
  theme_ipsum_rc(base_size = 18)+
  labs(title = "Qatar: R0",subtitle = "generation time (gamma distribution with a mean of 4 (SD=2)",x = "date")+
  geom_text_repel(mapping = aes(label = paste(dateRep,"\nR0 = ",round(R0,2),sep="")),size=2)#

Qatar_plot_daily <- ggplot(Qatar_cases_100,aes(dateRep,cases))+
  geom_hline(yintercept = 0)+
   geom_point(color="grey",size=3)+
  geom_line(color="grey",size=1.5)+
  theme_ipsum_rc(base_size = 18)+
  labs(title = "Qatar: daily cases",subtitle = "",x = "date")+
  geom_text_repel(mapping = aes(label = paste(dateRep,"\ncases = ",cases,sep="")),size=2)

```

<!-- START RMD display here -->

<span style="color:#0277BD;font-size: 18pt">data update from: `r max(data_in$dateRep)`</span>

# time line plots {.tabset .tabset-fade}

- **point saturation:** daily cases >> points more saturated = more daily cases

- **point size**: daily deaths >> points more saturated = more daily deaths

- start day = 10th case reported

- prediction based on data 7 days in the future


## Germany

```{r dailyGermany}
Germany_plot_daily
```

```{r RzeroGermany}
R0_Germany_plot
```


```{r GermanyTimeLine, echo=FALSE}

input_country <- "Germany"

tmp_country <- data_in %>% 
  filter(countriesAndTerritories==input_country)%>% 
  filter(day_sum_cases>=10)

plotly::ggplotly(
ggplot(data = tmp_country ,
       mapping = aes(day_start,
                   day_sum_cases,
                   label=dateRep))+
  geom_point(aes(size=deaths,alpha=cases))+
  stat_smooth(method = "gam",fullrange = T,formula = y ~ s(x, bs = "cs"),color="#01579B")+
  labs(size = "daily deaths",y = "total cases", x = "start of data recording (days) - since 31.12.2019",
       title = input_country)+
  xlim(c(min(tmp_country$day_start),max(tmp_country$day_start)+7))+
  theme_ipsum_rc(base_size = 13)
)
```

## USA

```{r dailyUSA}
USA_plot_daily
```


```{r RzeroUSA}
R0_USA_plot
```


```{r USATimeLine, echo=FALSE}

input_country <- "United_States_of_America"

tmp_country <- data_in %>% 
  filter(countriesAndTerritories==input_country)%>% 
  filter(day_sum_cases>=10)

plotly::ggplotly(
ggplot(data = tmp_country ,
       mapping = aes(day_start,
                   day_sum_cases,
                   label=dateRep))+
  geom_point(aes(size=deaths,alpha=cases))+
  stat_smooth(method = "gam",fullrange = T,formula = y ~ s(x, bs = "cs"),color="#01579B")+
  labs(size = "daily deaths",y = "total cases", x = "start of data recording (days) - since 31.12.2019",
       title = input_country)+
  xlim(c(min(tmp_country$day_start),max(tmp_country$day_start)+7))+
  theme_ipsum_rc(base_size = 13)
)
```

## Italy

```{r ItalyTimeLine, echo=FALSE}

input_country <- "Italy"

tmp_country <- data_in %>% 
  filter(countriesAndTerritories==input_country)%>% 
  filter(day_sum_cases>=10)

plotly::ggplotly(
ggplot(data = tmp_country ,
       mapping = aes(day_start,
                   day_sum_cases,
                   label=dateRep))+
  geom_point(aes(size=deaths,alpha=cases))+
  stat_smooth(method = "gam",fullrange = T,formula = y ~ s(x, bs = "cs"),color="#01579B")+
  labs(size = "daily deaths",y = "total cases", x = "start of data recording (days) - since 31.12.2019",
       title = input_country)+
  xlim(c(min(tmp_country$day_start),max(tmp_country$day_start)+7))+
  theme_ipsum_rc(base_size = 13)
)
```

## Spain

```{r SpainTimeLine, echo=FALSE}

input_country <- "Spain"

tmp_country <- data_in %>% 
  filter(countriesAndTerritories==input_country)%>% 
  filter(day_sum_cases>=10)

plotly::ggplotly(
ggplot(data = tmp_country ,
       mapping = aes(day_start,
                   day_sum_cases,
                   label=dateRep))+
  geom_point(aes(size=deaths,alpha=cases))+
  stat_smooth(method = "gam",fullrange = T,formula = y ~ s(x, bs = "cs"),color="#01579B")+
  labs(size = "daily deaths",y = "total cases", x = "start of data recording (days) - since 31.12.2019",
       title = input_country)+
  xlim(c(min(tmp_country$day_start),max(tmp_country$day_start)+7))+
  theme_ipsum_rc(base_size = 13)
)
```

## France

```{r FranceTimeLine, echo=FALSE}

input_country <- "France"

tmp_country <- data_in %>% 
  filter(countriesAndTerritories==input_country)%>% 
  filter(day_sum_cases>=10)

plotly::ggplotly(
ggplot(data = tmp_country ,
       mapping = aes(day_start,
                   day_sum_cases,
                   label=dateRep))+
  geom_point(aes(size=deaths,alpha=cases))+
  stat_smooth(method = "gam",fullrange = T,formula = y ~ s(x, bs = "cs"),color="#01579B")+
  labs(size = "daily deaths",y = "total cases", x = "start of data recording (days) - since 31.12.2019",
       title = input_country)+
  xlim(c(min(tmp_country$day_start),max(tmp_country$day_start)+7))+
  theme_ipsum_rc(base_size = 13)
)
```

## United Kingdom

```{r UKTimeLine, echo=FALSE}

input_country <- "United_Kingdom"

tmp_country <- data_in %>% 
  filter(countriesAndTerritories==input_country)%>% 
  filter(day_sum_cases>=10)

plotly::ggplotly(
ggplot(data = tmp_country ,
       mapping = aes(day_start,
                   day_sum_cases,
                   label=dateRep))+
  geom_point(aes(size=deaths,alpha=cases))+
  stat_smooth(method = "gam",fullrange = T,formula = y ~ s(x, bs = "cs"),color="#01579B")+
  labs(size = "daily deaths",y = "total cases", x = "start of data recording (days) - since 31.12.2019",
       title = input_country)+
  xlim(c(min(tmp_country$day_start),max(tmp_country$day_start)+7))+
  theme_ipsum_rc(base_size = 13)
)
```

## Sweden

```{r dailySweden}
Sweden_plot_daily
```


```{r RzeroSweden}
R0_Sweden_plot
```


```{r SwedenTimeLine, echo=FALSE}

input_country <- "Sweden"

tmp_country <- data_in %>% 
  filter(countriesAndTerritories==input_country)%>% 
  filter(day_sum_cases>=10)

plotly::ggplotly(
ggplot(data = tmp_country ,
       mapping = aes(day_start,
                   day_sum_cases,
                   label=dateRep))+
  geom_point(aes(size=deaths,alpha=cases))+
  stat_smooth(method = "gam",fullrange = T,formula = y ~ s(x, bs = "cs"),color="#01579B")+
  labs(size = "daily deaths",y = "total cases", x = "start of data recording (days) - since 31.12.2019",
       title = input_country)+
  xlim(c(min(tmp_country$day_start),max(tmp_country$day_start)+7))+
  theme_ipsum_rc(base_size = 13)
)
```


## Netherlands

```{r NetherlandsTimeLine, echo=FALSE}

input_country <- "Netherlands"

tmp_country <- data_in %>% 
  filter(countriesAndTerritories==input_country)%>% 
  filter(day_sum_cases>=10)

plotly::ggplotly(
ggplot(data = tmp_country ,
       mapping = aes(day_start,
                   day_sum_cases,
                   label=dateRep))+
  geom_point(aes(size=deaths,alpha=cases))+
  stat_smooth(method = "gam",fullrange = T,formula = y ~ s(x, bs = "cs"),color="#01579B")+
  labs(size = "daily deaths",y = "total cases", x = "start of data recording (days) - since 31.12.2019",
       title = input_country)+
  xlim(c(min(tmp_country$day_start),max(tmp_country$day_start)+7))+
  theme_ipsum_rc(base_size = 13)
)
```

## Switzerland

```{r SwitzerlandTimeLine, echo=FALSE}

input_country <- "Switzerland"

tmp_country <- data_in %>% 
  filter(countriesAndTerritories==input_country)%>% 
  filter(day_sum_cases>=10)

plotly::ggplotly(
ggplot(data = tmp_country ,
       mapping = aes(day_start,
                   day_sum_cases,
                   label=dateRep))+
  geom_point(aes(size=deaths,alpha=cases))+
  stat_smooth(method = "gam",fullrange = T,formula = y ~ s(x, bs = "cs"),color="#01579B")+
  labs(size = "daily deaths",y = "total cases", x = "start of data recording (days) - since 31.12.2019",
       title = input_country)+
  xlim(c(min(tmp_country$day_start),max(tmp_country$day_start)+7))+
  theme_ipsum_rc(base_size = 13)
)
```



## China

```{r ChinaTimeLine, echo=FALSE}

input_country <- "China"

tmp_country <- data_in %>% 
  filter(countriesAndTerritories==input_country)%>% 
  filter(day_sum_cases>=10)

plotly::ggplotly(
ggplot(data = tmp_country ,
       mapping = aes(day_start,
                   day_sum_cases,
                   label=dateRep))+
  geom_point(aes(size=deaths,alpha=cases))+
  stat_smooth(method = "gam",fullrange = T,formula = y ~ s(x, bs = "cs"),color="#01579B")+
  labs(size = "daily deaths",y = "total cases", x = "start of data recording (days) - since 31.12.2019",
       title = input_country)+
  xlim(c(min(tmp_country$day_start),max(tmp_country$day_start)+7))+
  theme_ipsum_rc(base_size = 13)
)
```


## South Korea

```{r SouthKoreaTimeLine, echo=FALSE}

input_country <- "South_Korea"

tmp_country <- data_in %>% 
  filter(countriesAndTerritories==input_country)%>% 
  filter(day_sum_cases>=10)

plotly::ggplotly(
ggplot(data = tmp_country ,
       mapping = aes(day_start,
                   day_sum_cases,
                   label=dateRep))+
  geom_point(aes(size=deaths,alpha=cases))+
  stat_smooth(method = "gam",fullrange = T,formula = y ~ s(x, bs = "cs"),color="#01579B")+
  labs(size = "daily deaths",y = "total cases", x = "start of data recording (days) - since 31.12.2019",
       title = input_country)+
  xlim(c(min(tmp_country$day_start),max(tmp_country$day_start)+7))+
  theme_ipsum_rc(base_size = 13)
)
```



## Qatar

```{r dailyQatar}
Qatar_plot_daily
```

```{r RzeroQatar}
R0_Qatar_plot
```


```{r QatarTimeLine, echo=FALSE}

input_country <- "Qatar"

tmp_country <- data_in %>% 
  filter(countriesAndTerritories==input_country)%>% 
  filter(day_sum_cases>=10)

plotly::ggplotly(
ggplot(data = tmp_country ,
       mapping = aes(day_start,
                   day_sum_cases,
                   label=dateRep))+
  geom_point(aes(size=deaths,alpha=cases))+
  stat_smooth(method = "gam",fullrange = T,formula = y ~ s(x, bs = "cs"),color="#01579B")+
  labs(size = "daily deaths",y = "total cases", x = "start of data recording (days) - since 31.12.2019",
       title = input_country)+
  xlim(c(min(tmp_country$day_start),max(tmp_country$day_start)+7))+
  theme_ipsum_rc(base_size = 13)
)
```



# current status 
```{r ValueBoxes,fig.width=10,fig.height=2}
SummaryValueBoxes
```

The cases/deaths reported since 2019-12-31 by country.

```{r stackedGlobalPlots,fig.width=10,fig.height=12}
global_cases_deaths_daily_plot/global_cases_deaths_overall_plot
```



```{r datatable}
DT::datatable(current_data_in, 
          rownames = F,
          filter = 'top', 
          options = list(pageLength = 10, autoWidth = TRUE),
          caption = 'cases/deaths reported since 2019-12-31 by country') %>% 
formatStyle(columns = c('total_deaths'),
  background = styleColorBar(range(unlist(current_data_in[,"total_deaths"]),na.rm = T), '#B71C1C'),#deaths
  backgroundSize = '98% 88%',
  backgroundRepeat = 'no-repeat',
  backgroundPosition = 'center') %>% 
formatStyle(columns = c('total_cases'),
  background = styleColorBar(range(unlist(current_data_in[,"total_cases"]),na.rm = T), '#01579B'),#cases
  backgroundSize = '98% 88%',
  backgroundRepeat = 'no-repeat',
  backgroundPosition = 'center') %>% 
formatStyle(columns = c('cases'),
  background = styleColorBar(range(unlist(current_data_in[,"cases"]),na.rm = T), '#006064'),#cases by day
  backgroundSize = '98% 88%',
  backgroundRepeat = 'no-repeat',
  backgroundPosition = 'center') %>% 
formatStyle(columns = c('deaths'),
  background = styleColorBar(range(unlist(current_data_in[,"deaths"]),na.rm = T), '#BF360C'),#deaths by day
  backgroundSize = '98% 88%',
  backgroundRepeat = 'no-repeat',
  backgroundPosition = 'center')
```

# cases and deaths over time {.tabset .tabset-fade}

The cases and deaths over time since 2019-12-31 by country. Top10 countries selected on individual total cases reported.


## reported cases aligned

cases over time aligned. Day 0 is the day were the 1000th case was reported.

```{r}
case_plot_1000th_case
```

```{r,fig.width=16,fig.height=10}
case_plot_1000th_case_facet
```


```{r,fig.width=12,fig.height=10}
cases_plot_1000th_case_per_day
```

## reported deaths aligned

deaths over time aligned. Day 0 is the day were the 5th death was reported.

```{r}
deaths_plot_1000th_case
```

```{r,fig.width=16,fig.height=10}
deaths_plot_1000th_case_facet
```


```{r,fig.width=12,fig.height=10}
deaths_plot_1000th_case_per_day
```



## reported cases

```{r}
cases_plot
```

## reported deaths

```{r}
deaths_plot
```




